%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Short Sectioned Assignment
% LaTeX Template
% Version 1.0 (5/5/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
\usepackage[english]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm} % Math packages

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont} % Make all sections centered, the default font and small caps

\usepackage{fancyhdr} % Custom headers and footers
\pagestyle{fancyplain} % Makes all pages in the document conform to the custom headers and footers
\fancyhead{} % No page header - if you want one, create it in the same way as the footers below
\fancyfoot[L]{} % Empty left footer
\fancyfoot[C]{} % Empty center footer
\fancyfoot[R]{\thepage} % Page numbering for right footer
\renewcommand{\headrulewidth}{0pt} % Remove header underlines
\renewcommand{\footrulewidth}{0pt} % Remove footer underlines
\setlength{\headheight}{13.6pt} % Customize the height of the header

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

\setlength\parindent{0pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height

\title{	
\normalfont \normalsize 
\textsc{University of Southern California} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge CS402 Assignment 01 -- Read/Write Locking \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

\author{Michael Gerow} % Your name

\date{\normalsize\today} % Today's date or a custom date

\begin{document}

\maketitle % Print the title

%----------------------------------------------------------------------------------------
%	PROBLEM 1
%----------------------------------------------------------------------------------------

\section{Read Write Locking}
Read write locking is implemented similarly to coarse grained locking in that we only have on global lock for the entire table, but instaed of using a pthread\_mutex\_t we instead use a pthread\_rwlock\_t. This allows us to use the functions pthread\_rwlock\_rdlock in order to get a read lock and pthread\_rwlock\_wrlock in order to get a write lock.

There is also one small difference with the way we compile under linux (which can be seen in the if (!initialized) clause within interpret\_command function).  The POSIX spec does not mentiona anything about giving readers or writers priority.  Because of that, a lot of the implementations seem to not really care about priority and just allow any waking thread to continue through the lock, as long as it is allowed.  This has the unfortunate effect of basically allowing readers to continue pouring through a lock even if a writer is waiting to acquire it.  In order to solve this we enable the attribute PTHREAD\_RWLOCK\_PREFER\_WRITER\_NONRECURSIVE\_NP when we are compiling under linux (which supports this attribute).  By doing this we should see an increase in performance when we have a relatively even number of readers and writers.

\subsection{Testing}
We test the correctness of the read write lock similarly to the corarse grained lock, using the test that adds a lot of elements to the database and then removes them, which should leave us with an empty database. This can, like with coarse, be done simply by running:

\begin{verbatim}
./server_rw < test/250_add_remove | tail -n 800 | less
\end{verbatim}

Of course, we should also expect an increase in speed for read write locking.  It turns out that this is only true for certain types of accesses of the database. Performance testing will be covered in the performance document.


\end{document}
